---
title: "P8105_hw2_gl2677"
author: "Gaotong LIU"
date: "9/26/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# install.packages("ggrepel")
library(tidyverse)
library(readxl)
library(ggrepel)
```

## Problem 0
#### Read and clean the Mr. Trash Wheel sheet
```{r}
mr_trash_wheel = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "Mr. Trash Wheel",
             range = "A2:N408") %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0),
         sports_balls = as.integer(sports_balls)) 

```

#### Read and clean precipitation data for 2017 and 2018
```{r}
prec_2017 = 
    read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2017 Precipitation",
             range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate(year = "2017",
         month = month.name) %>% 
  drop_na() 

prec_2018 = 
    read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2018 Precipitation",
             range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = "2018",
         month = month.name) %>% 
  drop_na() 

#Join 2017 and 2018 percipitation
prec_2017_2018 = bind_rows(prec_2017 , prec_2018) 
```

#### Interpretation
There are `r nrow(mr_trash_wheel)` observations and `r ncol(mr_trash_wheel)` variables in `mr_trash_wheel` dataset. This dataset has the information on the dumpter number, date of collection, amount of total litter (weight and volume), litter type and the number of homes powered by trash. In my opinion, the amount of total litter, litter type and homes powered are most important.
The key variables are ``r colnames(mr_trash_wheel[5:14])``. 
The median number of sports balls in a dumpster in 2017 was `r median(pull(filter(mr_trash_wheel, year == "2017"), sports_balls))`.

There are `r nrow(prec_2017_2018)` observations and `r ncol(prec_2017_2018)` variables in `prec_2017_2018` dataset. This dataset shows the total precipitation in the corresponding month and year.
The key variables are ``r colnames(prec_2017_2018[2])``. 
The total precipitation in 2018 was `r sum(pull(filter(prec_2017_2018, year == "2018"), total))`.

## Problem 2
#### Read and clean pols_month data
```{r , message=FALSE}
pols_month = 
  read_csv("./data/pols-month.csv") %>% 
  separate(mon, into = c("year","month","day"), sep = "-") %>% 
  mutate(year = as.numeric(year),
         month = as.numeric(month),
         month = month.name[month],
         president = factor(prez_dem, 
                            levels = c("gop"  = 0, "dem" = 1),
                            labels = c("gop","dem"))) %>% 
  select(-day, -prez_gop, -prez_dem)
```

#### Read and clean snp data
```{r message=FALSE}
snp = 
  read_csv("./data/snp.csv") %>% 
  separate(date, into = c("month","day","year"), sep = "/") %>% 
  mutate(year = as.numeric(year),
         month = as.numeric(month)) %>% 
    arrange(year,month) %>% 
  mutate(month = month.name[month]) %>% 
  select(year,month,close)
```

#### Read and clean unemployment data
```{r message=FALSE}
unemployment = 
  read_csv("./data/unemployment.csv") %>% 
  pivot_longer(Jan:Dec,
               names_to = "month",
               values_to = "percentage") %>% 
  rename(year = Year) %>% 
  mutate(month = match(month, month.abb),
         month = month.name[month])

pols_snp = left_join(pols_month, snp, by = c("year","month")) 
pols_snp_unemp = left_join(pols_snp, unemployment, by = c("year","month"))
```

#### Interpretation
There are `r nrow(pols_month)` observations and `r ncol(pols_month)` variables in `pols_month` dataset. This dataset shows the president's political party and the number of national politicians who are democratic or republican at any given time.
The key variables are ``r colnames(pols_month[3:9])``. The range of year is from `r min(pull(pols_month, year))` to `r max(pull(pols_month, year))`.

There are `r nrow(snp)` observations and `r ncol(snp)` variables in `snp` dataset. This dataset is related to Standard & Poor’s stock market index (S&P) at any given time.
The key variables are ``r colnames(snp[3])``. The range of year is from `r min(pull(snp, year))` to `r max(pull(snp, year))`.

There are `r nrow(unemployment)` observations and `r ncol(unemployment)` variables in `unemployment` dataset. This dataset shows unemployment percentage at any given time.
The key variables are ``r colnames(unemployment[3])``. The range of year is from `r min(pull(unemployment, year))` to `r max(pull(unemployment, year))`.

There are `r nrow(pols_snp_unemp)` observations and `r ncol(pols_snp_unemp)` variables in `pols_snp_unemp` **resulting dataset**. 
The key variables are ``r colnames(pols_snp_unemp[3:11])``. The range of year is from `r min(pull(pols_snp_unemp, year))` to `r max(pull(pols_snp_unemp, year))`.

## Problem 3
#### Read and clean baby_names data
```{r ,message=FALSE}
baby_names = 
  read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_lower(gender),
         ethnicity = str_to_lower(ethnicity),
         childs_first_name = str_to_lower(childs_first_name),
         ethnicity = recode(ethnicity, `asian and paci` = "asian and pacific islander", `black non hisp` = "black non hispanic", `white non hisp` = "white non hispanic")) %>% 
  distinct() 
```

#### Rank of olivia over time
```{r}
folivia = baby_names %>% 
  filter(childs_first_name == "olivia" & gender == "female") %>% 
  select(-childs_first_name, -gender, -count) %>% 
  pivot_wider(names_from = "year_of_birth",
              values_from = "rank") %>% 
  knitr::kable(caption = "Table 1. The rank in popularity of the name “Olivia” as a female baby name over time.")
folivia
```

#### Name of rank first over time
```{r}
mrank = baby_names %>% 
  filter(rank == 1 & gender == "male") %>% 
  select(-rank, -gender, -count) %>% 
  mutate(childs_first_name = str_to_title(childs_first_name)) %>% 
  pivot_wider(names_from = "year_of_birth",
              values_from = "childs_first_name") %>% 
  knitr::kable(caption = "Table 2. The most popular name among male children over time.")
mrank
```

#### Scatterplot
```{r}
baby_names %>% 
  filter(gender == "male" & ethnicity == "white non hispanic" & year_of_birth == "2016") %>% 
  mutate(childs_first_name = str_to_title(childs_first_name)) %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point(alpha = 0.5) + 
  geom_text_repel(aes(label = ifelse(rank <= 15, as.character(childs_first_name),'')),
                  box.padding   = 0.5,
                  point.padding = 0.5) +
  labs(title = "Figure 1. The number of children with a name against the rank in popularity ",
          subtitle = "for male, white non-hispanic children born in 2016") +
  theme(plot.subtitle = element_text(hjust = 0.5))

```